Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-01-29 paper.tex
lambda calculus and intro

Imagine an agent faced with a suite of new problems totally different
from anything it has seen before. It has at its disposal a basic set
of primitives it can compose to build solutions to these problems, but
it is no idea what kinds of primitives are appropriate for which
problems nor does it know the higher-level domain-specific language in
which solutions are best expressed.

@@

@@

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$z$};
  \node[latent] at (2.5,1.5) (tx){$\theta_x$};
  \node[obs] at (1,0) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$N$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node[fill = white]{$q(x)$} (tx.south);
  \draw [->,dashed] (tx.west) -- (z.east);
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $z$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta_x = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}
  \end{figure}

@@

\bibliography{example_paper}
\bibliographystyle{icml2018}

>>>>

Imagine an agent faced with a suite of new problems totally different
from anything it has seen before. It has at its disposal a basic set
of primitive actions it can compose to build solutions to these problems, but
it is no idea what kinds of primitives are appropriate for which
problems nor does it know the higher-level domain-specific language in
which solutions are best expressed.
How can our agent get off the ground?

The AI and machine learning literature contains two broad takes on this problem.
The first take is that the agent should come up with a better representation of the space of solutions,
for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill}.

@@

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$z$};
  \node[latent] at (2.5,1.5) (tx){$\theta_x$};
  \node[obs] at (1,0) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$N$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node[fill = white]{$q(x)$} (tx.south);
  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
  \draw [->,dashed] (tx.west) -- (z.east);
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $z$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta_x = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}
\end{figure}

\section{Program Representation}
We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
A $\lambda$-calculus expression is either:
\\\noindent A \emph{primitive}, like the number 5 or
  the function \texttt{sum}.
\\\noindent A \emph{variable}, like $x$, $y$, $z$
\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function. $\lambda$-abstractions have a variable and a body. The body is a $\lambda$-calculus expression. Abstractions are written as $\lambda \text{var}. \text{body}$.
\\\noindent An \emph{application} of a function to an argument. Both the function and the argument are $\lambda$-calculus expressions. The application of the function $f$ to the argument $x$ is written as $f\; x$.

For example, the function which squares the logarithm of a number is
$\lambda x.\text{\texttt{square}} (\text{\texttt{log} } x)$, and the identity function $f(x) = x$ is $\lambda x.x$. The
$\lambda$-calculus serves as a spartan but expressive Turing complete
program representation, and distills the essential features of functional languages like Lisp.

However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
(i.e., $5 \; (\lambda x.x)$).
We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
Hindley-Milner has two important features:
Feature 1: It supports \emph{parametric polymorphism}: meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$. A function that returns the first element of a list has the type $\texttt{list}(\alpha)\to\alpha$. Type variables are not the same has variables introduced by $\lambda$-abstractions.
Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
A detailed exposition of Hindley-Milner is beyond the scope of this work.

%% \begin{figure}
%%   \begin{align}
%%     \text{primitive types} & 
%%     \end{align}
%%   \end{figure}

@@

@@

\bibliography{main}
\bibliographystyle{icml2018}
